XCMS of round 1 and round 3 data combined for Toad Stage 22
========================================================
```{r echo=FALSE, results='hide', message=FALSE}
#source("http://bioconductor.org/biocLite.R")
#biocLite("multtest")
#biocLite("mzR")
#biocLite("faahKO")
#biocLite("preprocessCore")
#biocLite("clusterSim")
#biocLite("MetabolAnalyze")
#biocLite("mixOmics")
## try http:// if https:// URLs are not supported
source("https://bioconductor.org/biocLite.R")
biocLite("multtest")
library(xcms)
biocLite("faahKO")
biocLite("preprocessCore")
biocLite("clusterSim")
biocLite("MetabolAnalyze")
biocLite("mixOmics")
```

```{r echo=FALSE, results='hide', message=FALSE}
sys.frame(0)$ofile

#### read in and format data ####
if(Sys.info()[4]=="DC2626UTPURUCKE"){
metdir2<-"C://stp_drop//Dropbox//amphib_metabolomics//DATA//toad_stage22_1_3_p//"
cdf.path<-setwd(metdir2)
}

if(Sys.info()[4] == "stp-air.local"){
  metdir2<-path.expand("~/stp_drop/Dropbox/amphib_metabolomics/DATA/toad_stage22_1_3_p")
  cdf.path<-setwd(metdir2)
}
if(Sys.info()[4]=="Marcias-MacBook-Pro.local"){
  metdir2<-path.expand("~/Dropbox/amphib_metabolomics/DATA/toad_stage22_1_3_p")
  outdir2<-path.expand("~/Dropbox/amphib_metabolomics/DATA/toad_stage22_1_3_p")
  cdf.path<-setwd(metdir2)
}
file.exists(metdir2)
#### import cdf files 
#cdf.path<-setwd("~/Dropbox/amphib_metabolomics/DATA/toad_stage22_round1/polar/polar_toad1_cdf/xcms")
```
Temporal alignment with XCMS
```{r}
list.files(cdf.path)
cdffiles <- list.files(cdf.path, recursive = TRUE, full.names = TRUE)
cdffiles
# create xcms Set
x1 <- xcmsSet(cdffiles)
x1
#alignment across treatments and replicates
x2 <- group(x1) #match peaks across samples
#correct for retention time offsets
x3 <- retcor(x2, method="obiwarp", plottype = "deviation")
#x3 <- retcor(x2, family = "symmetric",peakwidth=c(20,50), plottype = "mdevden")
# group peaks again after correcting for time offset
x3 <- group(x3, bw = 10) 
#correct for retention time offsets
x4 <- retcor(x3, method = "obiwarp", plottype = "deviation")

# group peaks again after correcting for time offset
x4 <- group(x4, bw = 10) 
#fill peaks that were missed during first grouping
x4 <- fillPeaks(x3, method="chrom") #?
#normalize by peaks
x4_values<-groupval(x4, "medret", "into")
```
T-tests for control vs. 1250 
```{r}
out.diff1<-diffreport(x4, "control", "d1250", 10, metlin=FALSE, h=480, w=640)
group.c.1250sig<-NULL
for(i in 1:1687){
  group.c.1250sig[i] <- ifelse(out.diff1[i,4] < 0.05, 1,0)
}
group.c.sig.no<-sum(group.c.sig, na.rm=TRUE) #327
```

```{r}
write.csv(group1, "aligned_group1.csv")
write.csv(group2, "aligned_group3.csv")
```
 find z-scores for group1 vs. group3 t-tests for all dose levels
```{r}
#control group
group.1<-c(2,4,5,7,9)
group.2<-c(1,3,6,8,10)
group1<-x4_values[,group.1]
group2<-x4_values[,group.2]
group.c.p<-NULL
group.c.z<-NULL
for(i in 1:1678){
  t.out<-t.test(group1[i,], group2[i,])
  group.c.p[i]<-t.out$p.value
    group.c.z[i]<-t.out$statistic
}
group.c.sig<-NULL
for(i in 1:1687){
  group.c.sig[i] <- ifelse(group.c.p[i] < 0.05, 1,0)
}
group.c.sig.no<-sum(group.c.sig, na.rm=TRUE) #327 sig t-tests

# dose 10
group.1.10<-c(12,14,16,19)
group.2.10<-c(11,13,15,17,18)
group1.10<-x4_values[,group.1.10]
group2.10<-x4_values[,group.2.10]
group.10.p<-NULL
group.10.z<-NULL
for(i in 1:1678){
  t.out.10<-t.test(group1.10[i,], group2.10[i,])
  group.10.p[i]<-t.out.10$p.value
    group.10.z[i]<-t.out.10$statistic
}
group.10.sig<-NULL
for(i in 1:1687){
  group.10.sig[i] <- ifelse(group.10.p[i] < 0.05, 1,0)
}
group.10.sig.no<-sum(group.10.sig, na.rm=TRUE) #228 sig t-tests

# dose 50
group.1.50<-c(41,42,44,45,48)
group.2.50<-c(40,43,46,47)
group1.50<-x4_values[,group.1.50]
group2.50<-x4_values[,group.2.50]
group.50.p<-NULL
group.50.z<-NULL
for(i in 1:1678){
  t.out.50<-t.test(group1.50[i,], group2.50[i,])
  group.50.p[i]<-t.out.50$p.value
    group.50.z[i]<-t.out.50$statistic
}
group.50.sig<-NULL
for(i in 1:1687){
  group.50.sig[i] <- ifelse(group.50.p[i] < 0.05, 1,0)
}
group.50.sig.no<-sum(group.50.sig, na.rm=TRUE) # 784 sig t-tests

# dose 250
group.1.250<-c(30,32,34,37,38)
group.2.250<-c(31,33,35,36,39)
group1.250<-x4_values[,group.1.250]
group2.250<-x4_values[,group.2.250]
group.250.p<-NULL
group.250.z<-NULL
for(i in 1:1678){
  t.out.250<-t.test(group1.250[i,], group2.250[i,])
  group.250.p[i]<-t.out.250$p.value
    group.250.z[i]<-t.out.250$statistic
}
group.250.sig<-NULL
for(i in 1:1687){
  group.250.sig[i] <- ifelse(group.250.p[i] < 0.05, 1,0)
}
group.250.sig.no<-sum(group.250.sig, na.rm=TRUE) # 384 sig t-tests

# dose 1250
group.1.1250<-c(21,23,25,27,28)
group.2.1250<-c(20,22,24,26,29)
group1.1250<-x4_values[,group.1.1250]
group2.1250<-x4_values[,group.2.1250]
group.1250.p<-NULL
group.1250.z<-NULL
for(i in 1:1678){
  t.out.1250<-t.test(group1.1250[i,], group2.1250[i,])
  group.1250.p[i]<-t.out.1250$p.value
    group.1250.z[i]<-t.out.1250$statistic
}
group.1250.sig<-NULL
for(i in 1:1687){
  group.1250.sig[i] <- ifelse(group.1250.p[i] < 0.05, 1,0)
}
group.1250.sig.no<-sum(group.1250.sig, na.rm=TRUE) #704  sig t-tests
#create vector of sig t-tests for fig
grp1_2<-c(group.c.sig.no,group.10.sig.no,group.50.sig.no,group.250.sig.no,group.1250.sig.no)
```
t-tests to get false negative rate
```{r}
#### Controls #### 
v1<-c(1,2,3,4,5,6,7,8,9,10)
#t.test.results<- matrix(nrow=252,ncol=2)
ttest.array <- array(dim = c(1687,2,252)) #create array for results
sub_c<-x4_values[,1:10] #create df with just controls for sampling
#dim(t.test.results)
for(i in 1:252){
out1<-sample(v1,5, replace=FALSE)
temp1<-sub_c[,out1]
temp2<-sub_c[,-out1]
#temp1t<-t(temp1)
  for (j in 1:1687){
    t.out<-t.test(temp2[j,], temp1[j,])
    ttest.array[j,1,i]<-t.out$p.value
    ttest.array[j,2,i]<-t.out$statistic
  }
}
# add up significant ttests across tests
ttest.array2<- array(dim = c(1687,1,252)) #create array for results
for(i in 1:1687){
  for(j in 1:252){
        ttest.array2[i,1,j] <- ifelse(ttest.array[i,1,j] < 0.05, 1,0)
  }
}

# add up significant ttests 
ttest.c<-NULL
for (i in 1:252){
  ttest.c[i]<-(sum(ttest.array2[,1,i])) 
}
#ttest.c
mean.c<-mean(ttest.c) # mean = 60.2
par(mfrow=c(1,1))
hist(ttest.c)

# plot histogram of z-scores
#hist(ttest.array[1,2,])
for (i in 1:1687){
hist(ttest.array[i,2,])
}

for (i in 1:1687){
hist(ttest.array[i,2,])
}
```
Dose 10
```{r}
#### Dose 10 ####
choose(9,5)
#gives 126
v1<-c(11,12,13,14,15,16,17,18,19)
#t.test.results<- matrix(nrow=252,ncol=2)
ttest.array.10 <- array(dim = c(1687,2,126)) #create array for results
#dim(t.test.results)
for(i in 1:126){
out1<-sample(v1,5, replace=FALSE)
temp1<-x4_values[,out1]
temp2<-x4_values[,-out1]

  for (j in 1:1687){
    t.out<-t.test(temp2[j,], temp1[j,])
    ttest.array.10[j,1,i]<-t.out$p.value
    ttest.array.10[j,2,i]<-t.out$statistic
  }
}
# add up significant ttests across tests
ttest.array2.10<- array(dim = c(1687,1,126)) #create array for results
for(i in 1:1687){
  for(j in 1:126){
        ttest.array2.10[i,1,j] <- ifelse(ttest.array.10[i,1,j] < 0.05, 1,0)
  }
}

# add up significant ttests 
ttest.10<-NULL
for (i in 1:126){
  ttest.10[i]<-(sum(ttest.array2.10[,1,i])) 
}
#ttest.10
mean.10<-mean(ttest.10) # mean = 97.1
```
Dose 50
```{r}
choose(9,5)
#gives 126
v1<-c(40,41,42,43,44,45,46,47,48)
#t.test.results<- matrix(nrow=252,ncol=2)
ttest.array.50 <- array(dim = c(1687,2,126)) #create array for results
#dim(t.test.results)
for(i in 1:126){
out1<-sample(v1,5, replace=FALSE)
temp1<-x4_values[,out1]
temp2<-x4_values[,-out1]
  for (j in 1:1687){
    t.out.50<-t.test(temp2[j,], temp1[j,])
    ttest.array.50[j,1,i]<-t.out.50$p.value
    ttest.array.50[j,2,i]<-t.out.50$statistic
  }
}

# add up significant ttests across tests
ttest.array2.50<- array(dim = c(1687,1,126)) #create array for results
for(i in 1:1687){
  for(j in 1:126){
        ttest.array2.50[i,1,j] <- ifelse(ttest.array.50[i,1,j] < 0.05, 1,0)
  }
}

# add up significant ttests 
ttest.50<-NULL
for (i in 1:126){
  ttest.50[i]<-(sum(ttest.array2.50[,1,i])) 
}
#ttest.c
mean.50<-mean(ttest.50) # mean = 73
```
Dose 250
```{r}
choose(10,5)
#gives 252
v1<-c(30,31,32,33,34,35,36,37,38,39)
ttest.array.250 <- array(dim = c(1687,2,252)) #create array for results
#dim(t.test.results)
for(i in 1:252){
out1<-sample(v1,5, replace=FALSE)
temp1<-x4_values[,out1]
temp2<-x4_values[,-out1]
  for (j in 1:1687){
    t.out.250<-t.test(temp2[j,], temp1[j,])
    ttest.array.250[j,1,i]<-t.out.250$p.value
    ttest.array.250[j,2,i]<-t.out.250$statistic
  }
}

# add up significant ttests across tests
ttest.array2.250<- array(dim = c(1687,1,252)) #create array for results
for(i in 1:1687){
  for(j in 1:252){
        ttest.array2.250[i,1,j] <- ifelse(ttest.array.250[i,1,j] < 0.05, 1,0)
  }
}

# add up significant ttests 
ttest.250<-NULL
for (i in 1:252){
  ttest.250[i]<-(sum(ttest.array2.250[,1,i])) 
}
#ttest.c
mean.250<-mean(ttest.250) # mean = 104
```
Dose 1250
```{r}
choose(10,5)
#gives 252
v1<-c(20,21,22,23,24,25,26,27,28,29)
#t.test.results<- matrix(nrow=252,ncol=2)
ttest.array.1250 <- array(dim = c(1687,2,252)) #create array for results
#dim(t.test.results)
for(i in 1:252){
out1<-sample(v1,5, replace=FALSE)
temp1<-x4_values[,out1]
temp2<-x4_values[,-out1]
  for (j in 1:1687){
    t.out<-t.test(temp2[j,], temp1[j,])
    ttest.array.1250[j,1,i]<-t.out$p.value
    ttest.array.1250[j,2,i]<-t.out$statistic
  }
}

# add up significant ttests across tests
ttest.array2.1250<- array(dim = c(1687,1,252)) #create array for results
for(i in 1:1687){
  for(j in 1:252){
        ttest.array2.1250[i,1,j] <- ifelse(ttest.array.1250[i,1,j] < 0.05, 1,0)
  }
}
# add up significant ttests 
ttest.1250<-NULL
for (i in 1:252){
  ttest.1250[i]<-(sum(ttest.array2.1250[,1,i])) 
}
#ttest.c
mean.1250<-mean(ttest.1250) # mean = 499
```
Figure comparing group ttest difference with mean bootstrapped # sig ttests
```{r}
par(mfrow=c(1,1))
boot.mean.all<-c(mean.c, mean.10, mean.50, mean.250, mean.1250)
grp1_2
counts <- cbind(boot.mean.all, grp1_2)

bp<-barplot(as.matrix(t(counts)), main="Mean sig ttests per sample",
  xlab="Sig t-tests", col=c("lightblue","red"),
  legend = c("boot","actual"), beside=TRUE, axes=T, ylim=c(0,800))
text(bp, 0, round(counts, 1),cex=0.75,pos=3) 
```
Plot of z-scores for all samples
```{r}
#1250
plot(ttest.array.1250[,2,1], type='p', axes=T, ylim=c(-6,12),col=rgb(0,0,100,50,maxColorValue=255), xlab="Peak", ylab="t-stat", main="1250")
for (i in 1:252){
  points(ttest.array.1250[,2,i], type='p', col=rgb(0,0,100,50,maxColorValue=255))
}
points(group.1250.z, col="red")

#250
plot(ttest.array.250[,2,1], type='p', axes=T, ylim=c(-8,9),col=rgb(0,0,100,50,maxColorValue=255),xlab="Peak", ylab="t-stat", main="250" )
for (i in 1:252){
  points(ttest.array.250[,2,i], type='p', col=rgb(0,0,100,20,maxColorValue=255))
  }
points(group.250.z, col="red")

#50
plot(ttest.array.50[,2,1], type='p', axes=T, ylim=c(-16,12),col=rgb(0,0,100,50,maxColorValue=255),xlab="Peak", ylab="t-stat", main="50" )
for (i in 1:126){
  points(ttest.array.50[,2,i], type='p', col=rgb(0,0,100,50,maxColorValue=255))
  }
points(group.50.z, col="red")

#10
plot(ttest.array.10[,2,1], type='p', axes=T, ylim=c(-14,12),col=rgb(0,0,100,50,maxColorValue=255),xlab="Peak", ylab="t-stat", main="10" )
for (i in 1:126){
  points(ttest.array.10[,2,i], type='p', col=rgb(0,0,100,50,maxColorValue=255))
  }
points(group.10.z, col="red")

#control
plot(ttest.array[,2,1], type='p', axes=T, ylim=c(-20,12),col=rgb(0,0,100,50,maxColorValue=255),xlab="Peak", ylab="t-stat", main="control" )
for (i in 1:126){
  points(ttest.array[,2,i], type='p', col=rgb(0,0,100,50,maxColorValue=255))
  }
points(group.c.z, col="red")
```
Plot histogram of mean # of t-tests significant
```{r}
#fix mean and sd of pnorm
hist(ttest.c, breaks=40)
mean.c<-mean(ttest.c)
sd.c<-sd(ttest.c) #128.4
z.c<-(327-48) / sd.c #z.c=2.2
pnorm(z.c, mean=mean.c, sd=sd.c) #0.36

hist(ttest.10, breaks=40)
mean.10<-mean(ttest.10)
sd.10<-sd(ttest.10) #90.9
z.10<-(228-68) / sd.10 #z.10 = 1.8
pnorm(z.10, mean.10, sd.10) #0.23

hist(ttest.50, breaks=40)
mean.50<-mean(ttest.50)
sd.50<-sd(ttest.50) #91.2
z.50<-(784-79) / sd.50 #z.50=7.7
pnorm(z.50, mean=mean.50, sd=sd.50) #0.21

hist(ttest.250, breaks=40)
mean.250<-mean(ttest.250)
sd.250<-sd(ttest.250) #131.1
z.250<-(384-94) / sd.250 #z.250=2.2
pnorm(z.250, mean=mean.250, sd=sd.250) #0.24

hist(ttest.1250, breaks=40)
mean.1250<-mean(ttest.1250)
sd.1250<-sd(ttest.1250) #261.5
z.1250<-(704-467) / sd.1250 #z.1250=0.91
pnorm(z.1250, mean=mean.1250, sd=sd.1250) #0.037
```

```{r}
#control group
group.1<-c(2,4,5,7,9)
group.2<-c(1,3,6,8,10)
group1<-x4_values[,group.1]
group2<-x4_values[,group.2]
group.c.p<-NULL
group.c.z<-NULL
for(i in 1:1678){
  t.out<-t.test(group1[i,], group2[i,])
  group.c.p[i]<-t.out$p.value
    group.c.z[i]<-t.out$statistic
}
group.c.sig<-NULL
for(i in 1:1687){
  group.c.sig[i] <- ifelse(group.c.p[i] < 0.05, 1,0)
}
group.c.sig.no<-sum(group.c.sig, na.rm=TRUE) #327 sig t-tests

```
Compare distributions of abundance values between groups
```{r}
#controls
group.1<-c(2,4,5,7,9)
group.2<-c(1,3,6,8,10)
group1<-x4_values[,group.1]
group2<-x4_values[,group.2]
mean.group1<-NULL
mean.group2<-NULL
group1.25<-NULL
group1.75<-NULL
for( i in 1:1687){
  mean.group1[i]<-mean(group1[i,])
  mean.group2[i]<-mean(group2[i,])
  group1.25[i]<-quantile(group1[i,], c(0.25))
  group1.75[i]<-quantile(group1[i,], c(0.75))
}
plot(mean.group2, type="p", col="blue", pch=19, cex=0.5, ylim=c(0,500000))
#points(mean.group2, col="red")
points(group1.25, col="black", pch="-")
points(group1.75, col="black", pch="-")


# dose 10
group.1.10<-c(12,14,16,19)
group.2.10<-c(11,13,15,17,18)
group1.10<-x4_values[,group.1.10]
group2.10<-x4_values[,group.2.10]
mean.group1.10<-NULL
mean.group2.10<-NULL
for( i in 1:1687){
  mean.group1.10[i]<-mean(group1.10[i,])
  mean.group2.10[i]<-mean(group2.10[i,])
}
plot(mean.group1.10, type="p", col="blue", ylim=c(0,500000))
points(mean.group2.10, col="red")
# 50
group.1.50<-c(41,42,44,45,48)
group.2.50<-c(40,43,46,47)
group1.50<-x4_values[,group.1.50]
group2.50<-x4_values[,group.2.50]
mean.group1.50<-NULL
mean.group2.50<-NULL
for( i in 1:1687){
  mean.group1.50[i]<-mean(group1.50[i,])
  mean.group2.50[i]<-mean(group2.50[i,])
}
plot(mean.group1.50, type="p", col="blue", ylim=c(0,500000))
points(mean.group2.50, col="red")
#250
group.1.250<-c(30,32,34,37,38)
group.2.250<-c(31,33,35,36,39)
group1.250<-x4_values[,group.1.250]
group2.250<-x4_values[,group.2.250]
mean.group1.250<-NULL
mean.group2.250<-NULL
for( i in 1:1687){
  mean.group1.250[i]<-mean(group1.250[i,])
  mean.group2.250[i]<-mean(group2.250[i,])
}
plot(mean.group1.250, type="p", col="blue", ylim=c(0,500000))
points(mean.group2.250, col="red")
#1250
group.1.1250<-c(21,23,25,27,28)
group.2.1250<-c(20,22,24,26,29)
group1.1250<-x4_values[,group.1.1250]
group2.1250<-x4_values[,group.2.1250]
mean.group1.1250<-NULL
mean.group2.1250<-NULL
for( i in 1:1687){
  mean.group1.1250[i]<-mean(group1.1250[i,])
  mean.group2.1250[i]<-mean(group2.1250[i,])
}
plot(mean.group1.1250, type="p", col="blue", ylim=c(0,500000))
points(mean.group2.1250, col="red")
```
Normalize by sum of sample which converts to a sum of 1 for all peaks in a sample
```{r}
x4_norm<-data.frame(matrix(nrow=1687, ncol=48)) 
for(i in 1:48){
sum.temp<-sum(x4_values[,i])
  for (j in 1:1687){
    x4_norm[j,i]<-x4_values[j,i] / sum.temp
  }
} 
x4_norm.names<-colnames(x4_values)
colnames(x4_norm)<-x4_norm.names
write.csv(x4_norm, "x4_norm.csv")
plot(x4_norm[,5], type="l", lwd=2, col="blue", ylim=c(0,0.03))
lines(x4_norm[,6], col="red", lwd=.5)
```
Bootstrapped t-tests to look at false negative rate with NORMALIZED data
find z-scores for group1 vs. group2 t-tests for all dose levels
```{r}
#control group
group.1<-c(2,4,5,7,9)
group.2<-c(1,3,6,8,10)
group1<-x4_norm[,group.1]
group2<-x4_norm[,group.2]
group.c.p<-NULL
group.c.z<-NULL
for(i in 1:1678){
  t.out<-t.test(group1[i,], group2[i,])
  group.c.p[i]<-t.out$p.value
    group.c.z[i]<-t.out$statistic
}
group.c.sig<-NULL
for(i in 1:1687){
  group.c.sig[i] <- ifelse(group.c.p[i] < 0.05, 1,0)
}
group.c.sig.no<-sum(group.c.sig, na.rm=TRUE) #974 sig t-tests

# dose 10
plot(x4_norm[,12], type="l", lwd=2, col="blue", ylim=c(0,0.03))
lines(x4_norm[,11], col="red", lwd=.5)
group.1.10<-c(12,14,16,19)
group.2.10<-c(11,13,15,17,18)
group1.10<-x4_norm[,group.1.10]
group2.10<-x4_norm[,group.2.10]
group.10.p<-NULL
group.10.z<-NULL
for(i in 1:1678){
  t.out.10<-t.test(group1.10[i,], group2.10[i,])
  group.10.p[i]<-t.out.10$p.value
    group.10.z[i]<-t.out.10$statistic
}
group.10.sig<-NULL
for(i in 1:1687){
  group.10.sig[i] <- ifelse(group.10.p[i] < 0.05, 1,0)
}
group.10.sig.no<-sum(group.10.sig, na.rm=TRUE) #861 sig t-tests

# dose 50
group.1.50<-c(41,42,44,45,48)
group.2.50<-c(40,43,46,47)
group1.50<-x4_norm[,group.1.50]
group2.50<-x4_norm[,group.2.50]
group.50.p<-NULL
group.50.z<-NULL
for(i in 1:1678){
  t.out.50<-t.test(group1.50[i,], group2.50[i,])
  group.50.p[i]<-t.out.50$p.value
    group.50.z[i]<-t.out.50$statistic
}
group.50.sig<-NULL
for(i in 1:1687){
  group.50.sig[i] <- ifelse(group.50.p[i] < 0.05, 1,0)
}
group.50.sig.no<-sum(group.50.sig, na.rm=TRUE) # 926 sig t-tests

# dose 250
group.1.250<-c(30,32,34,37,38)
group.2.250<-c(31,33,35,36,39)
group1.250<-x4_norm[,group.1.250]
group2.250<-x4_norm[,group.2.250]
group.250.p<-NULL
group.250.z<-NULL
for(i in 1:1678){
  t.out.250<-t.test(group1.250[i,], group2.250[i,])
  group.250.p[i]<-t.out.250$p.value
    group.250.z[i]<-t.out.250$statistic
}
group.250.sig<-NULL
for(i in 1:1687){
  group.250.sig[i] <- ifelse(group.250.p[i] < 0.05, 1,0)
}
group.250.sig.no<-sum(group.250.sig, na.rm=TRUE) # 908 sig t-tests

# dose 1250
group.1.1250<-c(21,23,25,27,28)
group.2.1250<-c(20,22,24,26,29)
group1.1250<-x4_norm[,group.1.1250]
group2.1250<-x4_norm[,group.2.1250]
group.1250.p<-NULL
group.1250.z<-NULL
for(i in 1:1678){
  t.out.1250<-t.test(group1.1250[i,], group2.1250[i,])
  group.1250.p[i]<-t.out.1250$p.value
    group.1250.z[i]<-t.out.1250$statistic
}
group.1250.sig<-NULL
for(i in 1:1687){
  group.1250.sig[i] <- ifelse(group.1250.p[i] < 0.05, 1,0)
}
group.1250.sig.no<-sum(group.1250.sig, na.rm=TRUE) #949  sig t-tests
#create vector of sig t-tests for fig
grp1_2<-c(group.c.sig.no,group.10.sig.no,group.50.sig.no,group.250.sig.no,group.1250.sig.no)
```
t-tests to get false negative rate for NORMALIZED data
```{r}
#### Controls #### 
v1<-c(1,2,3,4,5,6,7,8,9,10)
#t.test.results<- matrix(nrow=252,ncol=2)
ttest.array <- array(dim = c(1687,2,252)) #create array for results
sub_c<-x4_norm[,1:10] #create df with just controls for sampling
#dim(t.test.results)
for(i in 1:252){
out1<-sample(v1,5, replace=FALSE)
temp1<-sub_c[,out1]
temp2<-sub_c[,-out1]
#temp1t<-t(temp1)
  for (j in 1:1687){
    t.out<-t.test(temp2[j,], temp1[j,])
    ttest.array[j,1,i]<-t.out$p.value
    ttest.array[j,2,i]<-t.out$statistic
  }
}
# add up significant ttests across tests
ttest.array2<- array(dim = c(1687,1,252)) #create array for results
for(i in 1:1687){
  for(j in 1:252){
        ttest.array2[i,1,j] <- ifelse(ttest.array[i,1,j] < 0.05, 1,0)
  }
}

# add up significant ttests 
ttest.c<-NULL
for (i in 1:252){
  ttest.c[i]<-(sum(ttest.array2[,1,i])) 
}
#ttest.c
mean.c<-mean(ttest.c) # mean = 76
par(mfrow=c(1,1))
hist(ttest.c)

# plot histogram of z-scores
#hist(ttest.array[1,2,])
for (i in 1:1687){
hist(ttest.array[i,2,])
}

for (i in 1:1687){
hist(ttest.array[i,2,])
}
```
Dose 10
```{r}
#### Dose 10 ####
choose(9,5)
#gives 126
v1<-c(11,12,13,14,15,16,17,18,19)
#t.test.results<- matrix(nrow=252,ncol=2)
ttest.array.10 <- array(dim = c(1687,2,126)) #create array for results
#dim(t.test.results)
for(i in 1:126){
out1<-sample(v1,5, replace=FALSE)
temp1<-x4_norm[,out1]
temp2<-x4_norm[,-out1]

  for (j in 1:1687){
    t.out<-t.test(temp2[j,], temp1[j,])
    ttest.array.10[j,1,i]<-t.out$p.value
    ttest.array.10[j,2,i]<-t.out$statistic
  }
}
# add up significant ttests across tests
ttest.array2.10<- array(dim = c(1687,1,126)) #create array for results
for(i in 1:1687){
  for(j in 1:126){
        ttest.array2.10[i,1,j] <- ifelse(ttest.array.10[i,1,j] < 0.05, 1,0)
  }
}

# add up significant ttests 
ttest.10<-NULL
for (i in 1:126){
  ttest.10[i]<-(sum(ttest.array2.10[,1,i])) 
}
#ttest.10
mean.10<-mean(ttest.10) # mean = 104
```

Dose 50
```{r}
choose(9,5)
#gives 126
v1<-c(40,41,42,43,44,45,46,47,48)
#t.test.results<- matrix(nrow=252,ncol=2)
ttest.array.50 <- array(dim = c(1687,2,126)) #create array for results
#dim(t.test.results)
for(i in 1:126){
out1<-sample(v1,5, replace=FALSE)
temp1<-x4_norm[,out1]
temp2<-x4_norm[,-out1]
  for (j in 1:1687){
    t.out.50<-t.test(temp2[j,], temp1[j,])
    ttest.array.50[j,1,i]<-t.out.50$p.value
    ttest.array.50[j,2,i]<-t.out.50$statistic
  }
}

# add up significant ttests across tests
ttest.array2.50<- array(dim = c(1687,1,126)) #create array for results
for(i in 1:1687){
  for(j in 1:126){
        ttest.array2.50[i,1,j] <- ifelse(ttest.array.50[i,1,j] < 0.05, 1,0)
  }
}

# add up significant ttests 
ttest.50<-NULL
for (i in 1:126){
  ttest.50[i]<-(sum(ttest.array2.50[,1,i])) 
}
#ttest.c
mean.50<-mean(ttest.50) # mean = 111
```
Dose 250
```{r}
choose(10,5)
#gives 252
v1<-c(30,31,32,33,34,35,36,37,38,39)
ttest.array.250 <- array(dim = c(1687,2,252)) #create array for results
#dim(t.test.results)
for(i in 1:252){
out1<-sample(v1,5, replace=FALSE)
temp1<-x4_norm[,out1]
temp2<-x4_norm[,-out1]
  for (j in 1:1687){
    t.out.250<-t.test(temp2[j,], temp1[j,])
    ttest.array.250[j,1,i]<-t.out.250$p.value
    ttest.array.250[j,2,i]<-t.out.250$statistic
  }
}

# add up significant ttests across tests
ttest.array2.250<- array(dim = c(1687,1,252)) #create array for results
for(i in 1:1687){
  for(j in 1:252){
        ttest.array2.250[i,1,j] <- ifelse(ttest.array.250[i,1,j] < 0.05, 1,0)
  }
}

# add up significant ttests 
ttest.250<-NULL
for (i in 1:252){
  ttest.250[i]<-(sum(ttest.array2.250[,1,i])) 
}
#ttest.c
mean.250<-mean(ttest.250) # mean = 69
```

Dose 1250
```{r}
choose(10,5)
#gives 252
v1<-c(20,21,22,23,24,25,26,27,28,29)
#t.test.results<- matrix(nrow=252,ncol=2)
ttest.array.1250 <- array(dim = c(1687,2,252)) #create array for results
#dim(t.test.results)
for(i in 1:252){
out1<-sample(v1,5, replace=FALSE)
temp1<-x4_norm[,out1]
temp2<-x4_norm[,-out1]
  for (j in 1:1687){
    t.out<-t.test(temp2[j,], temp1[j,])
    ttest.array.1250[j,1,i]<-t.out$p.value
    ttest.array.1250[j,2,i]<-t.out$statistic
  }
}

# add up significant ttests across tests
ttest.array2.1250<- array(dim = c(1687,1,252)) #create array for results
for(i in 1:1687){
  for(j in 1:252){
        ttest.array2.1250[i,1,j] <- ifelse(ttest.array.1250[i,1,j] < 0.05, 1,0)
  }
}
# add up significant ttests 
ttest.1250<-NULL
for (i in 1:252){
  ttest.1250[i]<-(sum(ttest.array2.1250[,1,i])) 
}
#ttest.c
mean.1250<-mean(ttest.1250) # mean = 152

grp1_2<-c(group.c.sig.no,group.10.sig.no,group.50.sig.no,group.250.sig.no,group.1250.sig.no)
```
Figure comparing group ttest difference with mean bootstrapped # sig ttests
```{r}
par(mfrow=c(1,1))
boot.mean.all<-c(mean.c, mean.10, mean.50, mean.250, mean.1250)
grp1_2
counts <- cbind(boot.mean.all, grp1_2)

bp<-barplot(as.matrix(t(counts)), main="Mean sig ttests per sample",
  xlab="Sig t-tests", col=c("lightblue","red"),
  legend = c("boot","actual"), beside=TRUE, axes=T, ylim=c(0,800))
text(bp, 0, round(counts, 2),cex=0.75,pos=1) 
```

Create giant matrix for giant regression with group and dose as factors
```{r}

View(x4_norm)
df.norm<-as.data.frame(x4_norm)
View(df.norm)
df.norm.t<-t(df.norm)
View(df.norm.t)
group<-c("B","A","B","A","A","B","A","B","A","B",
         "B","A","B","A","B","A","B","B","A","B",
         "A","B","A","B","A","B","A","A","B","A",
         "B","A","B","A","B","B","A","A","B","B",
         "A","A","B","A","A","B","B","A")
trt<-c("C","C","C","C","C","C","C","C","C","C","d10","d10","d10","d10","d10","d10","d10","d10","d10","d1250","d1250","d1250","d1250","d1250","d1250","d1250","d1250","d1250","d1250","d250","d250","d250","d250","d250","d250","d250","d250","d250","d250","d50","d50","d50","d50","d50","d50","d50","d50","d50")
df.norm.t2<-cbind(group, df.norm.t)
View(df.norm.t2)
df.norm.t3<-cbind(trt,df.norm.t2)
View(df.norm.t3)
sample_id<-(colnames(x4_values))
df.norm.t4<-cbind(sample_id,df.norm.t3)
View(df.norm.t4)
df.norm.t4<-as.data.frame(df.norm.t4)
df.norm.t4$sample_id<-as.factor(df.norm.t4$sample_id)
# add column names for peaks
peaks<-rownames(x4_values)
colnames(df.norm.t4)[4:1690]<-peaks
View(df.norm.t4)
md <- melt(df.norm.t4, id=(c("sample_id", "group", "trt")))
View(md)
write.csv(md, "group1_group3_long.csv")
```
Regression with giant matrix:
 names(md)
 "sample_id" "group"     "trt"       "variable"  "value"   

```{r}
data_control<-md[,which(md$trt=="C" | md$trt=="d10")]                  
data_control_d10<-subset(md, md$trt=="C" | md$trt=="d10")                  
View(data_control_d10)
data_control_d10$value<-as.numeric(as.character(data_control_d10$value))
summary(data_control_d10)
unique_peaks<-unique(md$variable)
vector <- character(length(unique_peaks))
for(peak in unique_peaks){
  temp_data<-data_control_d10[which(data_control_d10$variable==peak),]                  
  temp_out<-lm(value~group+trt, data=temp_data)
  append(vector, temp_out$statistic[3])
  }
summary(temp_out)
```














